Final Project - Stroke Predictions

### 2021-05-13 draft.Rmd
data viz

---
title: "Important Factors in Predicting a Stroke"
author: "Group 10"
output: 
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

@@

# Exploratory Data Analysis
•	Exploratory Data Analysis – Initial summary statistics and graphs with an emphasis on variables you believe to be important for your analysis.

@@

-which columns have missing values?

@@

```{r, echo = FALSE}
# changing variable types for analysis and replacing missing values
stroke$bmi <- gsub("N/A", NA, stroke$bmi)
stroke$smoking_status <- gsub("Unknown", NA, stroke$smoking_status)
stroke$gender <- as.factor(stroke$gender)
stroke$bmi <- as.numeric(stroke$bmi)
stroke$age <- as.numeric(stroke$age)
stroke$avg_glucose_level <- as.numeric(stroke$avg_glucose_level)
stroke$ever_married <- as.factor(stroke$ever_married)
stroke$work_type <- as.factor(stroke$work_type)
stroke$Residence_type <- as.factor(stroke$Residence_type)
stroke$smoking_status <- as.factor(stroke$smoking_status)

summary(stroke)
```

@@

The two variables were are most interested in are not very balanced. On the left is a histogram for hypertension, and on the right is one for heart disease.As mentioned before, only 9.75% of the patients having hypertension and 5.40% of them having heart disease

@@

```{r, echo = FALSE}
# plotting the distribution of bmi based on smoking status

@@

# strip plots?
```

@@

This is a visualization of the correlation plot, which confirms that no variables appear to be highly correlated with one another. The dark blue diagonal circles are just the variables correlated with themselves.

@@

@@

>>>>

---
title: "Final Project - Stroke Predictions"
author: "Group 10"
output: 
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

@@

# Exploratory Data Analysis

@@

@@

```{r, echo = FALSE}
# changing variable types for analysis and replacing missing values
stroke$bmi <- gsub("N/A", NA, stroke$bmi)
stroke$smoking_status <- gsub("Unknown", NA, stroke$smoking_status) #replacing unknown with NA is smoking status
stroke$gender <- as.factor(stroke$gender)
stroke$bmi <- as.numeric(stroke$bmi)
stroke$age <- as.numeric(stroke$age)
stroke$avg_glucose_level <- as.numeric(stroke$avg_glucose_level)
stroke$ever_married <- as.factor(stroke$ever_married)
stroke$work_type <- as.factor(stroke$work_type)
stroke$Residence_type <- as.factor(stroke$Residence_type)
stroke$smoking_status <- as.factor(stroke$smoking_status)
stroke$hypertension <- as.factor(stroke$hypertension)
stroke$heart_disease <- as.factor(stroke$heart_disease)
stroke$stroke <- as.factor(stroke$stroke)

summary(stroke)

@@

The two variables we are most interested in are not very balanced. On the left is a histogram for hypertension, and on the right is one for heart disease. As mentioned before, only 9.75% of the patients have hypertension and 5.40% of them have heart disease. This could lead to issues in creating predictive models later.

@@

Next we wanted to look at some visualizations of numeric variables we think will be important in stroke predictions, other than the important binary variables like hypertension and heart disease.

```{r, echo = FALSE, message = FALSE,warning = FALSE}
# plotting the distribution of bmi based on smoking status

@@

#plotting smoking status vs glucose level?

# strip plots?

# plotting bmi with stroke incidence
ggplot(stroke, 
       aes(y = stroke,
           x = bmi,
           color = stroke)) +
   geom_jitter(alpha = 0.2,
              width = .1) + 
  geom_boxplot(size = 1,
               outlier.shape = 1,
               outlier.color = "black",
               outlier.size = 3,
               color="black",
               alpha=0.2) +
  labs(title = "BMI distribution with Stroke Incidence")
```

The bmi vs. stroke incidence chart shows that the average bmi for patients who had a stroke was higher than those who didn't.

Diabetes is also another risk factor for stroke. Although this dataset does not have presence of diabetes as a variable, it does have average glucose level, which can be used to classify if patients are diabetic. The American Diabetes Association classifies normal blood glucose levels as less than 100 mg/dl, prediabetes is 100 mg/dl to 125 mg/dl, and diabetes is 126 mg/dl or higher. Looking at a plot of average blood glucose levels

```{r,echo = FALSE, message = FALSE}
#plotting avg. glucose level with stroke incidence
ggplot(stroke, 
       aes(y = stroke,
           x = avg_glucose_level,
           color = stroke)) +
   geom_jitter(alpha = 0.2,
              width = .1) + 
  geom_boxplot(size = 1,
               outlier.shape = 1,
               outlier.color = "black",
               outlier.size = 3,
               color="black",
               alpha=0.2) +
  labs(title = " Average Glucose Level Distribution with Stroke Incidence")
```

This graph doesn't give quite as clear of a result as the previous one, mainly because the distribution of the data is a little strange. There's a lot of people with somewhat normal levels and then a lot of people with high levels, but not as many in between. The average glucose level was still higher for patients who had a stroke.

Another numeric variable that would be interestng to look at would be age, as the risk for stroke increases with age.

```{r, echo = FALSE, message = FALSE}
#plotting age vs. stroke incidence
ggplot(stroke, 
       aes(y = stroke,
           x = age,
           color = stroke)) +
   geom_jitter(alpha = 0.2,
              width = .1) +

  labs(title = "Age Distribution with Stroke Incidence")
```

The scatterplot cleary indicates that those patients who had strokes were older, 40+ years on average.

As confirmed by some of the research we did, the most at-risk patient would be someone with hypertension and heart-disease, with a high bmi and average glucose level, who smokes, and who is older.

@@

This is a visualization of the correlation plot, which confirms that no variables appear to be highly correlated with one another. There is small correlations between age and hypertension, heart disease, high glucose levels, and stroke incidence. Some of the common comorbid conditions also have smaller correlation levels, like heart disease, stroke, hypertension, and average glucose level.

@@

-the dataset was kind of unbalanced

@@

https://www.diabetes.org/a1c/diagnosis
